On Extensibility and Personalizability of Multi-Modal Trip Planning

### 2016-03-03 paper.tex
Spell check.

\section{Introduction}
Trip planning, as an application of planning and scheduling,
has seen substansive implementations by researchers and
developers\cite{bast2015route}.
Some of the planning systems are \tit{multi-modal}; that is,
combining distinct transportation modes, the trip planners
compute optimal routes from sources to destinations.
This notion of ``optimality" generally refers to the computed routes
having minimal total time or total fare.

However, in the eyes of a user, it may be more faceted than just
``fastest" or ``cheapest."
For instance, for a college student who specifies that she will
only walk or take public transit in a trip from Palo Alto to
San Francisco, the computed plan is not necessarily the fastest
(e.g., taking a cab could be faster)
or the cheapest (e.g., walking all the way has no fare).
This happens when a user tells the planner her hard constraints,
called \tit{constraints}.  
The planner then needs to either satisfy the 
constraints in the search process or return failure because they are
over-restrictive.
Moreover, the user might want to further customize the planner by
describing soft constraints, called \tit{preferences}.
For example, an agent wants to travel from school to
downtown, and prefers biking to taking the bus.
Thus, a trip with more biking than bus may be considered better for
the agent than the one with more bus than biking.
In this case, the planner will need to accommodate user preferences
whenever possilbe in the search of optimal solutions.

@@

Our paper is organized as follows.
In the next section, we present what it means for a planner to be extensible
and formally define the method to incorporating new metrics into the planner.
In the following section, we discuss the two aspects of personalization
in trip planning: constraints and preferences, and how they are modeled and
reasoned with in the setting of multi-modal trip planning.
We then move on to describe the system structure of our graph-search based planner,
and show results obtained from our planner in various occassions.
Finally, we conclude by outlining some future research directions.

>>>>

\section{Introduction}
Trip planning, as an application of planning and scheduling,
has seen substantive implementations by researchers and
developers\cite{bast2015route}.
Some of the planning systems are \tit{multi-modal}; that is,
combining distinct transportation modes, the trip planners
compute optimal routes from sources to destinations.
This notion of ``optimality" generally refers to the computed routes
having minimal total time or total fare.

However, in the eyes of a user, it may be more faceted than just
``fastest" or ``cheapest."
For instance, for a college student who specifies that she will
only walk or take public transit in a trip from Palo Alto to
San Francisco, the computed plan is not necessarily the fastest
(e.g., taking a cab could be faster)
or the cheapest (e.g., walking all the way has no fare).
This happens when a user tells the planner her hard constraints,
called \tit{constraints}.  
The planner then needs to either satisfy the 
constraints in the search process or return failure because they are
over-restrictive.
Moreover, the user might want to further customize the planner by
describing soft constraints, called \tit{preferences}.
For example, an agent wants to travel from school to
downtown, and prefers biking to taking the bus.
Thus, a trip with more biking than bus may be considered better for
the agent than the one with more bus than biking.
In this case, the planner will need to accommodate user preferences
whenever possible in the search of optimal solutions.

@@

Our paper is organized as follows.
In the next section, we present what it means for a planner to be extensible
and formally define the method to incorporating new metrics into the planner.
In the following section, we discuss the two aspects of personalization
in trip planning: constraints and preferences, and how they are modeled and
reasoned with in the setting of multi-modal trip planning.
We then move on to describe the system structure of our graph-search based planner,
and show results obtained from our planner in various occasions.
Finally, we conclude by outlining some future research directions.
